---
title: "Forecast U.S.A Elections 2020"
author: "Yingrui"
date: "November 02, 2020"
output:
  pdf_document: default
  pdf_html: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Title of your Report
Forecast U.S.A Elections 2020

# Name(s) of Author(s) 
Yingrui

# Date
November 02, 2020

## Introduction
Forecasting presidential elections in the USA has become a meaningful endeavor owing to the fact that the country plays a significant role in the world markets. Every election period in the country does not only attract national but also international interest as the country's currency is the driver of exchange rate around the globe.In this case, through understanding the performance and opinions of the general public, it would be possible to determine the candidate that is most likely to win the elections. This report is based on evaluating and presenting a forecast on the candidate that will win the 2020 elections.  

## Data

The utilised Dataset were taken from two types of sources. First, the individual level data with public opinion was based on Democracy Fund + UCLA Nationscape full data from the a recent month of July 25, 2020. 

On the other hand,the other utilised data was based on a national survey on the posibilities of a candidate winning the elections. This data is based on the ACS (American Community Surveys) post stratification data.The data was used as gathered from the IPSUM website in credit of (Ruggles et al., 2020).

Based on the nature of the Analysis, it is critical to understand the contributions of each dataset in the work. Therefore, A breakdown of the cleaned is explained below.

## Democracy Fund and UCLA Nationscape (Full DataSet)

This dataset mainly provides the census data of the USA population. It is critical to the analysis as it elaborates on the voter eligible voters that are present in the country. The proposed model will be consider this data against survey to determine the candidate that has the highest popular vote. 
Based on the provided population in the census, the person with the least age had 2years whilst the one with the highest age had 97years in the assessed sample population. Based on the assessment of the population, it is evident that there more people eligible people to vote between the age of 18 to 97 as compared to non-voters who are between the age of 2 to 17 years. These results are presented in the pie chart below.

```{r}
slices <- c(110651, 496356)
lbls <- c("Age 2 to 17", "Age 18 to 97")

pie(slices, labels = lbls, main="Eligibility to Vote")
```
Figure 1: Eligibility to vote.

The aim of doing the assessment is to establish that the census data was viable to use in the prediction of the candidate that was likely to win in terms of the conducted survey. 

## Post-stratification Data

This dataset provided more information about the survey on the likelihood of each participating candidate to win the upcoming election. The considered variables in this case were the "vote_2020" and "Registration". In essence, the registration variable was important as it represented the number of registered voters that participated in the survey. On the other hand, the vote_2020 was important as it shows the tally between the candidates in the elections "Joe Biden" and "Donald Trump". 
On the other hand, the variable "interest" was considered for the purpose of assessing the amount of registered voters who had active interest to participate in the process. 

```{r}
library(janitor)
library(tidyverse)

pop <- read_csv("C:/Users/rmjes/OneDrive/Desktop/All Files/forecasting_us_election/outputs/population.csv")

mytable <- table(pop)

lbls <- paste(names(mytable), "\n", mytable, sep="")
pie(mytable, labels = lbls,
   main="Pie Chart of Voters 2020\n (with sample sizes)")
```
Figure 2: Descriptive Representation of Survey

Overall, based on the evaluation of the survey results from the data, it is evident that most of the population in the survey showed interest in electing Joe Biden as opposed to DOnald Trump. Nonetheless, as earlier indicated the results cannot be solely dependent on this tabulation given that there are other critical factors to note such as the "Registered" variable that represents eligible voters in the elections and "interest" variable that represents willingness of voters to participate in the elections. 

## Model

For the accurate prediction of the candidate that is likely to win in the election, a multiple linear regression was implemented. This model is utilised as it helps with evaluating previous predictions and forecasting the possible outcome based on the performance of variables (Gelman and Hill, 2006). 

In this case, to make an accurate prediction, it was critical to evaluate the variables that influence winning possibilities in the survey. In this case, the important variables that influence outcome of the elections were "Interest", which represents person's intention to vote "Registered", represents the number of registered voters in sample population, and "Vote_2020 that represents opinions of people based on the election candidates.

To use the Linear regression model, it was important to first run the regression for year 2020 and compare it with previous elections from 2016. A summary of the 2016 Elections was as follows:

```{r}
library(janitor)
library(tidyverse)

pop <- read_csv("C:/Users/rmjes/OneDrive/Desktop/All Files/forecasting_us_election/outputs/2016.csv")

mytab <- table(pop)

lbls <- paste(names(mytab), "\n", mytab, sep="")
pie(mytab, labels = lbls,
   main="Pie Chart of Voters 2016\n (with sample sizes)")

```
In that case, the popular contestants were Hilary Clinton and Donald Trump. In this case, Donald Trump had the highest vote of 2186 as compared to 2021 for Hilary Clinton. 

In this case, the utilized model was as follows: 
Y = B0 + B1X1 + B2X2 +…+B8X8 + e

Where "Y" is to be the predicted value. The X1 and X2 would represent the coefficients of the of the independent variables mentioned above. As it is a linear equation, the main assumption is that there was an existing linear relationship between the variables (Ismail et al., 2009). In achieving the prediction value, the coefficients are multiplied with the derived population to ensure that a standard figure is acquired. 

```{r}
library(randomForest)
Elect_Results_1 <- read.csv(
              "C:/Users/rmjes/OneDrive/Desktop/All Files/forecasting_us_election/outputs/population.csv",
              sep=",",
              na.strings = "",
              blank.lines.skip = TRUE,
                stringsAsFactors=TRUE)
```
## Mode


```{r}
target <- Elect_Results_1

newdata <- data.frame(
  candidate1= "Donald Trump",
  candidate2= "Joe Biden"
)

rf1 <- randomForest(target,data=Elect_Results_1,ntree=1000,proximity=TRUE)

```


## References
Gelman, A. and Hill, J. (2006) Data Analysis Using Regression and Multilevel/Hierarchical Models, Data Analysis Using Regression and Multilevel/Hierarchical Models. doi: 10.1017/cbo9780511790942.

Ismail, Zuhaimy et al. (2009) ‘Forecasting Gold Prices Using Multiple Linear Regression Method’, American Journal of Applied Sciences, 6(8), pp. 1509–1514.

Ruggles, S. et al. (2020) ‘IPSUM USA data extract’. Minneapolis, MN: IPUMS USA, (PUMS USA: Version 10.0 [dataset]). doi: https://doi.org/10.18128/D010.V10.0.

